A Bayesian networks approach to infer social changes in burials in northeastern Taiwan during the European colonization period

### 2021-06-04 analysis/paper/paper.Rmd
edit the main text

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  dpi = 300,
  fig.path = "../figures/"
)

image_width_1_col_inch = 90/25.4 # image Single column	w = 90 mm 
image_width_2_col_inch = 190/25.4 # image Double column w = 190 mm 
image_width_3_col_inch = 190/19
```

@@

<!---network analysis approaches and basic concepts-->
Network analysis has been increasingly used by archaeologists to understand past relationships, interactions, and structures of observed phenomena by visualizing and analyzing relational data [@Brandes2013; @Mills2017; @Peeples2019]. Recent developments in modeling approaches enable hypothesis testing for the intercorrelations between individual elements and overall structures in networks [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010]. By comparing hypothesized networks with the observed network, statistical network modeling can answer anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way. Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulations for network patterns [@Harris2013; @Handcock2008; @Ghafouri2020]. ERGMs are promising for evaluating the dynamic social processes behind observed archaeological networks [@Brughmans2014]. However, computational difficulties, sensitivities to uncertainties, and ambiguities in interpretation limit the practical applications of ERGMs [@Caimo2014].

@@

<!--Our burial network and research implication-->
Burial studies are usually based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining and comparing those individual observations to infer the organization of past societies [@Byrd1995; @Seikel2011]. Recent studies on burials using network analysis demonstrates it is a promising tool to test expectations derived from hypotheses [@Sosna2013]. Here we use network modeling to explore similarities and differences between burials since it provides a statistical way for identifying underlying variable structuring burial relations. We treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e. individual burials), pair-wise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013]. Our network indicates status similarity between individuals, where a tie means two individuals sharing a similar social status or a similar social environment that higher status individuals will be given similar burial goods [eg. @Collar2015, pp. 4; eg. @Sosna2013]. This is because burial goods, especially high value ones, normally convey cultural symbol of personal wealth or social status from which we can infer social differentiation [@Gamble2002; @Janes2013]. Our case study illustrates how a Bayesian approach to network modeling provides new insights into the formation of archaeological networks and allows testing of anthropological models to better understand dynamic social processes.

<!-- intro to networks: move to SI -->
<!-- two types of network analysis: move to SI -->

@@

<!-- current approaches to network analysis: move to SI -->

<!-- Bayesian inference for Exponential Random Graph Models -->

In this paper, we used Bayesian inference on ERGMs to test our hypotheses. ERGMs are the tool that provides the conceptual background for network formation and structure, allowing testing social processes through covariates. The benefit of the Bayesian part is to deal with computational issues of ERGMs mentioned above. Bayesian approaches to ERGMs are effective because it incorporates prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and uncertainty [@Caimo2017; @Lehmann2020]. Prior information can be derived from previous data or assumptions derived from the context of our data. An important advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011]. With the exchange algorithm, Bayesian ERGMs avoid doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives better convergence results. This enables us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].

By fitting an ERGM with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information to estimate the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019]. This allows robust and interpretable uncertainty quantification by examining the posterior mean and 95% credible intervals [@Caimo2017bayesian]. In addition, Bayesian approaches are useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies. @Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can accurately predict, on average, 80% of the ties when a third of data is missing.

@@

<!--archaeological specific case study and our model-->
We use a Bayesian ERGM to study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan (Figure \@ref(fig:figure-KWL-map)), which was occupied from the 14th to 19th centuries. This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007]. Our anthropological model proposes that the influence of an indirect colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994]. Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2004; @Cheng2017]. We assume that the use of imported goods and their values were increased and stressed after the European arrival. Different from Heping dao and Tamsui in northern Taiwan where European forts located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities due to local natural resources, such as rice and deerskins [@Chen2004; @Cheng2017]. This makes northeastern Taiwan a special area to explore the effect of European colonization in a pericolonial context that is still not well understood compared to European colonial bases [@Acabado2017; @Trabert2017].

The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017]. Network-based societies are identified in some places in Asia that were involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006]. The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by European colonial activities in northern Taiwan. Foreign trade goods could be used as prestige goods based on their archaeological contexts where very few of trade goods were found in middens and more in burials, along with a symbol of personal influence described in the Spanish historical records [@LiandWu2006]. In addition, the study of spatial distribution of trade ornaments in residential area at Kiwulan demonstrates a more clustered pattern after the European presence, indicative of intentional accumulation of trade goods [@LWandBM2020ornament].

<!--theoretical framework and research questions-->
The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000]. These are contrasted with corporate-based societies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age-based or kin-based, rather than individuals [@Siegel1999]. The European contact and associated trade activities could be the stimulus of competition for foreign resources among aggrandizers that led to social changes. Using the theoretical framework of corporate-network strategies, we test a hypothesis of the shift in organizational modes in relation to increased social inequality, from a more corporate to a more network society, after the European presence that can be observed in the Kiwulan burial network. We ask: (1) Do the burial network after European arrival present a more centralized structure with fewer subgroups compared to the burial network before European contact? and if so, (2) what are the major variables affecting or forming these two networks? Can we identify the shift in variables from ritual to wealth that associated with two different strategies? By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.

@@

```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/000-prep-data.R"))}
suppressWarnings(source(here::here("analysis", "paper", "000-prep-data.R")))
```

```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/001-data-tidy.R"))}
suppressWarnings(source(here::here("analysis", "paper", "001-data-tidy.R")))
```

<!-- Kiwulan burials and a review of current studies of social organization -->
We analyzed burial data collected from the excavation reports, and the original fieldwork notes for the upper component of Kiwulan (1350-1850 AD) [@Chen2007]. A total of 90 burials were unearthed from adjacent excavation squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)). The preservation of human remains at Kiwulan was generally poor, so we have incomplete age and sex data (details in Supplementary Online Materials) and lack health status data. Burials are oriented in an east-west direction on the north side of the residential area, which is indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement of houses. Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences. For example, @Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society. However, @Hsieh2012 suggested a relatively egalitarian structure based on a comparative analysis of the frequencies of all burial goods. She found that the burials with rare prestige goods were usually associated with elders, which might indicate achieved, rather than inherited, status. One important limitation of these previous studies is that they did not use analytical units suitable for comparing behaviors before and after European arrival. Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period (before European colonization) to the post-European period (during and after European colonization until arrival of Han Chinese immigrants).

```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/002-burial-location.R"))}
suppressWarnings(source(here::here("analysis", "paper", "002-burial-location.R")))
```

@@

<!-- Our chronology and principles of network building -->
To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`). Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [for more details see @LWandBM2020ornament]. We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size. We also excluded `r disturb` burials that were heavily disturbed by modern construction which prevented accurate determination of their chronology. Differences in network size were evaluated using vertex bootstrap in a later section.

Based on the assumption that social status and associated relations can be represented by sharing similar prestige goods [cf. @Coward2013, pp. 252], we built networks where burials (nodes in the network) that are linked when they have the same prestige goods in common. The prestige goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, and fish-shaped ornaments. These items are considered as high-value across different archaeological contexts throughout Taiwan based on their rarity and descriptions in historical records [@Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011]. European historical accounts mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006]. For example, a Spanish priest described a Spanish soldier exchanging carnelian beads for natural resources with local Indigenous people, because of the beads’ high value in Indigenous culture [@LiandWu2006]. In addition, Spanish visitors observed that Indigenous people possessing more imported goods were recognized as having higher status in their community [@LiandWu2006]. Besides, those goods are made of durable materials that is reasonable to assume that they were less affected by taphonomic processes.

@@

```{r bergm-pre, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/003-burials-pre-network-modelling.R"))}
suppressWarnings(source(here::here("analysis", "paper", "003-burials-pre-network-modelling.R")))
```

```{r bergm-post, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/004-burials-post-network-modelling.R"))}
suppressWarnings(source(here::here("analysis", "paper", "004-burials-post-network-modelling.R")))
```

@@

```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/005-network-diagrams-two-phases.R"))}
suppressWarnings(source(here::here("analysis", "paper", "005-network-diagrams-two-phases.R")))
```

@@

```{r bergm-prior-and-iteration}
prior_mean <- as_tibble(cbind(pre_prior_mean, 
                              post_prior_mean)) %>% 
  mutate(mean_same = ifelse(pre_prior_mean == post_prior_mean, pre_prior_mean, NA))

prior_sigma <- as_tibble(cbind(diag(pre_prior_sigma), 
                               diag(post_prior_sigma))) %>% 
  rename(pre_prior_sigma = V1, post_prior_sigma = V2) %>% 
  mutate(sigma_same = ifelse(pre_prior_sigma == post_prior_sigma, pre_prior_sigma, NA))

@@

```{r posterior-distribution, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/006-bergm-distribution-stats.R"))}
suppressWarnings(source(here::here("analysis", "paper", "006-bergm-distribution-stats.R")))
```

@@

```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/paper/007-bootstrap-CI.R"))}
suppressWarnings(source(here::here("analysis", "paper", "007-bootstrap-CI.R")))
```

>>>>

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  dpi = 300,
  fig.path = "../figures/"
)
```

@@

<!--network analysis approaches and basic concepts-->
Network analysis has been increasingly used by archaeologists to understand past relationships, interactions, and structures of observed phenomena by visualizing and analyzing relational data [@Brandes2013; @Mills2017; @Peeples2019]. Recent developments in modeling approaches enable hypothesis testing for the intercorrelations between individual elements and overall structures in networks [@Brughmans2013; @Brughmans2018; @Freeman2004; @Salvini2010]. By comparing hypothesized networks with the observed network, statistical network modeling can answer anthropological questions related to exchange [@Gjesfjeld2013; @Crabtree2015], diffusion [@Ostborn2015], or social transformation [@Mills2013] in a statistical way. Among network modeling methods, exponential random graph models (ERGMs) are stochastic models for investigating the process of network formation through dependence assumptions for relationships and simulations for network patterns [@Harris2013; @Handcock2008; @Ghafouri2020]. ERGMs are promising for evaluating the dynamic social processes behind observed archaeological networks [@Brughmans2014]. However, computational difficulties, sensitivities to uncertainties, and ambiguities in interpretation limit the practical applications of ERGMs [@Caimo2014].

@@

<!--Our burial network and research implication-->
Burial studies are usually based on characterizing each physical trace or burial variable, such as biological records, grave forms and goods, or ritual behaviors, and finally combining and comparing those individual observations to infer the organization of past societies [@Byrd1995; @Seikel2011]. Recent studies on burials using network analysis demonstrates it is a promising tool to test expectations derived from hypotheses [@Sosna2013]. Here we use network modeling to explore similarities and differences between burials by identifying underlying variable structuring burial relations. We treat burials as a complex network based on the assumption that a network is a patterned aggregation that includes individual elements (i.e. individual burials), pairwise relationships (the dyads, for example burials with similar types and amounts of grave goods), and an overall structure showing social patterns represented in the data [@Brandes2013]. Our network indicates status similarity between individuals, where a tie means two individuals sharing a similar social status or a similar social environment that higher status individuals will be given similar burial goods [eg. @Collar2015, pp. 4; eg. @Sosna2013]. This is because burial goods, especially high value ones, normally convey cultural symbol of personal wealth or social status from which we can infer social differentiation [@Gamble2002; @Janes2013]. Our case study illustrates how a Bayesian approach to network modeling provides new insights into the formation of archaeological networks and allows testing of anthropological models to better understand dynamic social processes.

<!--intro to networks: moved to the SI-->
<!--two types of network analysis: moved to the SI-->

@@

<!--current approaches to network analysis: move to the SI-->

<!--Bayesian inference on Exponential Random Graph Models--> 
In this paper, we used Bayesian inference on ERGMs to test our hypotheses. ERGMs are used as the tool to provide the conceptual background for network formation and structure, allowing testing social processes through covariates. The benefit of the Bayesian part is to deal with computational issues of ERGMs by incorporating prior information about the network configurations (details in Supplementary Online Materials) to better understand dependencies of network variables and uncertainty [@Caimo2017; @Lehmann2021]. Prior information can be derived from previous data or assumptions based on the context of data. An important advantage that Bayesian modeling has over traditional ERGMs is the application of Markov chain Monte Carlo (MCMC) simulation using the approximate exchange algorithm [@Caimo2011]. With the exchange algorithm, Bayesian ERGMs avoid doubly-intractable computations by directly sampling from the not normalized part of the posterior, which alleviates the computational problems and gives better convergence results. This enables us to deal with complicated dependence patterns with ease, providing better estimations for complex social network models with heterogeneous data [@Caimo2017; @Snijders2006].

By fitting an ERGM with the approximate exchange algorithm, a Bayesian approach generates posterior probabilities that incorporate our sample data and prior information to estimate the effect of each ERGM parameter in our models [@Caimo2015; @Nemmers2019]. This allows robust and interpretable uncertainty quantification by examining the posterior distribution [@Caimo2017bayesian]. In addition, Bayesian approaches are useful to deal with missing data, which is often a problem leading to misinterpretation of networks, especially for archaeological studies. @Koskinen2010 shows that the effect of missing data can be reduced with Bayesian modeling that can accurately predict, on average, 80% of the ties when a third of data is missing.

@@

<!--Intro to our case study and model-->
We use a Bayesian ERGM to study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan (Figure \@ref(fig:figure-KWL-map)), which was occupied from the 14th to 19th centuries. This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally an influx of Han Chinese in the 19th century [@Chen2007]. Our anthropological model proposes that the influence of an indirect colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994]. Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2005; @Cheng2017]. We assume that the use of imported goods and their values were amplified after the European arrival. Different from Heping dao and Tamsui in northern Taiwan where European forts are located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities due to local natural resources, such as rice and deerskins [@Chen2005; @Cheng2017]. This makes northeastern Taiwan a special area to explore the effect of European colonization in a pericolonial context that is still not well understood compared to European colonial bases [@Acabado2017; @Trabert2017].

<!--theoretical framework: corporate/network strategies-->
The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign goods [@Blanton1996; @Feinman2000; @Klehm2017]. The features of network-based societies include the accumulation of prestige goods, wealth differentiation, and trade monopolization through individual networks [@Blanton1996; @Feinman2000]. These are contrasted with corporate-based societies that stress communal ritual elements, shared power, and where wealth differentiation, if any, would be associated with corporate groups, such as age-based or kin-based, rather than individuals [@Siegel1999]. Network-based societies are identified in some places in Asia that were involved in long-distance trade [@Carter2015; @Ueda2016; @Liu2006].

<!--archaeological context of Kiwulan-->
The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by European colonial activities in northern Taiwan. Foreign trade goods were believed to be used as prestige goods based on archaeological contexts at Kiwulan [@Hsieh2009; @Wang2011]. There are very few trade goods found in middens but more in residential area and burials, along with a symbol of personal influence described in the Spanish historical records [@LiandWu2006]. In addition, the study of spatial distribution of trade ornaments in Kiwulan residential area demonstrates a more clustered pattern after the European presence, indicative of intentional accumulation of trade goods [@LWandBM2020ornament]. The European contact and associated trade activities could be the stimulus of competition for foreign resources among aggrandizers that led to social changes.

<!--hypothesis and research questions-->
Using the theoretical framework of corporate-network strategies, we test a hypothesis of the shift in organizational modes in relation to increased social inequality, from a more corporate to a more network society, after the European presence that can be observed in the Kiwulan burial network. We ask: (1) Do the burial network after European arrival present a more centralized structure with fewer subgroups compared to the burial network before European contact? and if so, (2) what are the major variables affecting or forming these two networks? Can we identify the shift in variables from ritual to wealth associated with two different strategies? By answering these questions, this case study helps to expand our understanding of European colonial effects on Indigenous groups in a pericolonial context.

@@

```{r import-and-inspect-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/000-prep-data.R"))}
suppressWarnings(source(here::here("analysis", "code", "000-prep-data.R")))
```

```{r tidy-data, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/001-data-tidy.R"))}
suppressWarnings(source(here::here("analysis", "code", "001-data-tidy.R")))
```

<!-- Kiwulan burials and a review of current studies of social organization -->
We analyzed burial data collected from the excavation reports, and the original fieldwork notes for the upper component of Kiwulan (1350-1850 AD) [@Chen2007]. A total of 90 burials were unearthed from adjacent excavation squares that provide continuous stratigraphic sections suitable for temporal comparison (Figure \@ref(fig:figure-burial-map)). The preservation of human remains at Kiwulan was generally poor, so we have incomplete data of age and sex (details in Supplementary Online Materials) and lack health status data. Burials are oriented in an east-west direction on the north side of the residential area, which is indicated by post-holes and _in-situ_ wooden posts, suggesting a well-organized spatial arrangement of houses. Previous studies report an uneven distribution of prestige goods across burials without agreement about whether this uneven distribution hints at vertical social differences. For example, @Cheng2008 interpreted the unequal distribution of glass beads, especially the gold-foil beads between burials, as evidence for hierarchy, indicating a stratified society. However, @Hsieh2012 suggested a relatively egalitarian structure based on a comparative analysis of the frequencies of all burial goods. She found that the burials with rare prestige goods were usually associated with elders, which might indicate achieved, rather than inherited, status. One important limitation of these previous studies is that they did not use analytical units suitable for comparing behaviors from different phases. Here, we adopt a new chronological framework for the burials to test if network configurations differ from the pre-European period (before European colonization) to the post-European period (during and after European colonization until the arrival of Han Chinese immigrants).

```{r burial-location, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/002-burial-location.R"))}
suppressWarnings(source(here::here("analysis", "code", "002-burial-location.R")))
```

@@

<!-- Our chronology and principles of network building -->
To compare burial networks, we assigned burials to the pre-European period (n = `r pre_E`), European and post-European period (n = `r post_E`). Our assignments are based on an established fine-grained chronology that was reexamined and cross-validated by diagnostic materials, stratigraphic data, depth, and radiocarbon ages [for more details see @LWandBM2020ornament; @LWandBM2020pot]. We excluded burials from the Chinese phase (n = `r chi`) due to the smaller sample size. We also excluded `r disturb` burials that were heavily disturbed by modern construction which prevented accurate determination of their chronology. Differences in network size were evaluated using vertex bootstrap in a later section.

Based on the assumption that social status and associated relations can be represented by sharing similar prestige goods [cf. @Coward2013, pp. 252], we built networks where burials (nodes in the network) are linked when they have the same trade goods in common. The trade goods we identified include gold-foil beads, carnelian beads, glass beads, Chinese porcelains, stonewares, gold foils, fish-shaped ornaments, and imported coins, which were also less affected by taphonomic processes because of durable materials. These items are considered as prestige goods due to their rarity in burial contexts and descriptions in historical records [@Chen2007; @Cheng2008; @Hsieh2009; @Hsieh2012; @Wang2011]. European historical accounts mentioned that those items were treated as prestige goods in Indigenous culture [@Borao2009; @LiandWu2006]. For example, a Spanish priest described a Spanish soldier exchanging carnelian beads for natural resources with local Indigenous people, because of the beads’ high value in Indigenous culture [@LiandWu2006]. In addition, Spanish visitors observed that Indigenous people possessing more imported goods were recognized as having higher status in their community [@LiandWu2006].

@@

```{r custom-function, include=FALSE}
suppressWarnings(source(here::here("analysis", "code", "999-bgof-custom-function.R")))
```

```{r bergm-pre, include=FALSE}
suppressWarnings(source(here::here("analysis", "code", "003-burials-pre-network-modelling.R")))
```

```{r bergm-post, include=FALSE}
suppressWarnings(source(here::here("analysis", "code", "004-burials-post-network-modelling.R")))
```

@@

```{r network-diagrams, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/005-network-diagrams-two-phases.R"))}
suppressWarnings(source(here::here("analysis", "code", "005-network-diagrams-two-phases.R")))
```

@@

```{r bergm-prior-and-iteration}
prior_mean<- data.frame(cbind(pre_prior_mean, 
                              post_prior_mean)) %>% 
  mutate(mean_same = ifelse(pre_prior_mean == post_prior_mean, pre_prior_mean, NA))

prior_sigma <- data.frame(cbind(diag(pre_prior_sigma), 
                                diag(post_prior_sigma))) %>% 
  rename(pre_prior_sigma = X1, post_prior_sigma = X2) %>% 
  mutate(sigma_same = ifelse(pre_prior_sigma == post_prior_sigma, pre_prior_sigma, NA))

@@

```{r posterior-distribution, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/006-bergm-distribution-stats.R"))}
suppressWarnings(source(here::here("analysis", "code", "006-bergm-distribution-stats.R")))
```

@@

```{r bootstrap-CI-test, include=FALSE, cache=TRUE, cache.extra=tools::md5sum(here::here("analysis/code/007-bootstrap-CI.R"))}
suppressWarnings(source(here::here("analysis", "code", "007-bootstrap-CI.R")))
```
